---
title: "Assignment 8: Ensemble Methods"
author: "Talha Gulzar"
date: "11/17/2024"
output:
  rmdformats::readthedown:
    number_sections: true
    highlight: tango
    df_print: paged
    center: true
---

```{r setup, include=FALSE}
rm(list = ls()) # clears global environment
knitr::opts_knit$set(root.dir = 'C:/Users/LENOVO/OneDrive - Institute of Business Administration/Desktop/QTM6300')
```


# Context

Let’s take a look at the data located in UniversalBank.csv. Each row represents a customer at small but rapidly growing bank. The columns measure all sorts of customer characteristics, ranging from their demographic information (e.g., Age, Family) to whether they currently have various accounts open with the bank (e.g., Securities Account, CD Account). For a complete description of the fields, consult the data page on Canvas.

The bank is aggressively trying to convert customers from depositors into borrowers through its personal loan program. The column **Personal Loan** shows whether each customer responded to a direct marketing campaign related to this program. The marketing team is now trying to understand what types of customers responds to new personal loans marketing. If they can establish reasonably strong predictive power, they will deploy the model more widely across their customer base to identify promising leads and a more nuanced target market.


## Question 1

Do the following:

* Import the data as "bank"

* Remove any predictor that might be inappropriate for this activity.

* Set the target to logical.

```{r}
bank = read.csv("UniversalBank.csv")
bank$Personal.Loan = as.logical(bank$Personal.Loan)
bank = na.omit(bank)
bank$ID = NULL
bank$ZIP.Code = NULL
factor_columns = c("Education","Family","Securities.Account","Online","CreditCard","CD.Account")
bank[factor_columns] = lapply(bank[factor_columns], factor)
```


## Question 2

* Set a seed of 109 and partition a training set with 72% of the data.

* Run a Classification Tree to predict Personal.Loan, make the predictions, and calculate the error rate. What is the error rate?

Answer: 

```{r}
set.seed(109)
N = nrow(bank)
trainingSize  = round(N*0.72)
trainingCases = sample(N, trainingSize)
training  = bank[trainingCases,]
test      = bank[-trainingCases,]

library(rpart)
model = rpart(Personal.Loan ~ ., data=training)
pred = predict(model, test)
pred = (pred > 0.5)
library(rpart.plot)
rpart.plot(model)
error_tree = sum(pred!=test$Personal.Loan)/nrow(test)
error_tree


```


## Question 3

Bagging: Run a random forest model where you set the number of trees to 1000. Make the predictions and calculate the error rate. What is the error rate? Does this improve on the original single Classification Tree model?

Answer: 

```{r}

library(randomForest)


# Run the model, where ntree is equal to number of trees you want. set this large for first iteration!
rf = randomForest(Personal.Loan ~ ., data= training, ntree=1000)
pred_rf = predict(rf, test)
pred_rf = (pred_rf > 0.5)

error_randomforest = sum(pred_rf!=test$Personal.Loan)/nrow(test)
error_randomforest
# Do our errors decrease? 

```



## Question 4

* Boosting: Set a new seed to 23 and run a boosted tree model with 500 trees and 5 folds. Determine the best tree size. What is this size? Then use this "best tree size" to run another boosted tree model.

* Make the predictions and calculate the error rate. What is the error rate? If this error rate is wose than one or more of your previous models, what do you think might have happened?

Answer: 


```{r}
set.seed(23)

training$Personal.Loan=as.integer(training$Personal.Loan) # The gbm algorithm requires a target that is integer (0,1)
library(gbm)
boost = gbm(Personal.Loan ~ ., data=training,n.trees=500, cv.folds=5)

# Figure out what the best # of trees should be
# black is train; green is cross-val
best_size <- gbm.perf(boost,method="cv")
gbm.perf(boost)

# Run the model again with best tree size
boost = gbm(Personal.Loan ~ ., data=training,n.trees=500, cv.folds=5)


pred_boost  = predict(boost, test, n.trees=best_size, type="response")
pred_boost = (pred_boost > 0.5)

#calculate errors
error_boost = sum(pred_boost != test$Personal.Loan)/nrow(test)
error_boost
```


## Question 5

* Now, run a logistic regression model to predict Personal.Loan. Refine the model using the step() function.
* Make the predictions. Use a cutoff of 0.5. 
* Calculate the error rate. What is the error rate?

Answer: 


```{r}

training$Personal.Loan <- as.logical(training$Personal.Loan)

set.seed(109)

traininglg <- bank[trainingCases,]
testlg <- bank[-trainingCases,]

modellg <- glm(Personal.Loan ~ ., data=traininglg, family=binomial)
summary(modellg)

options(scipen=100)

modellg2 <- step(modellg)
summary(modellg2)

# store probabilities.   type=response (gives probabilities)
testlg$predictions <- predict(modellg2, testlg, type="response")
predictions <- testlg$predictions
#store TRUE/Falses 
testlg$predictionsTF <- (testlg$predictions >= 0.5)
predictionsTF <- testlg$predictionsTF

# store observed values too
observations <-testlg$Personal.Loan

#Error Rate
error_ratelg <- sum(predictionsTF != observations)/nrow(testlg)
error_ratelg

```


# Question 6

* Now, try to do stacking by combining the  Random Forest model, the Boosted model, and the logistic regression model. What is your error rate?

Answer: 

```{r}
# We need the predictions from multiple models that we have already run

# First, get the predictions for all of data frame observations, not just test.
pred_rf_full = predict(rf, bank)
pred_rf_full = (pred_rf_full > 0.5)

pred_boost_full = predict(boost, bank, n.trees=best_size, type="response")
pred_boost_full = (pred_boost_full > 0.5)

pred_log_full <- predict(modellg2, bank, type="response")
pred_log_full <- (pred_log_full > 0.5)

#Column bind the two data sets with the different predictions

bank_stacked = cbind(bank,pred_boost_full, pred_rf_full, pred_log_full)

bank_stacked$Personal.Loan = as.logical(bank$Personal.Loan)

# Set the training and test data; now these sets have the predictions!
train_stacked = bank_stacked[trainingCases, ]
test_stacked = bank_stacked[-trainingCases, ]

# Run the stacked algorithm, where the manager/meta model is a logistic model. 
stacked = glm(Personal.Loan ~ ., data = train_stacked, family=binomial)

# Make the predictions
pred_stacked = predict(stacked, test_stacked, type="response")
pred_stacked = (pred_stacked > 0.5)

# Calculate error rate
error_stacked = sum(pred_stacked != test$Personal.Loan)/nrow(test)
error_stacked
```



## Question 7

Which ensemble method lowered the error the most? 

Answer:With an error rate of 0.01214286, the random forest and stacked models had the lowest error rates among the models we ran in this worksheet.

