---
title: "Assignment 5: kNN Regression and Classification"
author: "Talha Gulzar"
date: "October 8, 2024"
format: html
editor: source
self-contained: true
toc: true
toc-expand: true
toc-depth: 3
---


```{r setup, include=FALSE}
rm(list = ls()) # clear working directory
knitr::opts_knit$set(root.dir = 'C:/Users/USER/OneDrive - Institute of Business Administration/Desktop/QTM 666')
```

# Context

Let’s take a look at the data located in UniversalBank.csv. Each row represents a customer at small but rapidly growing bank. The columns measure all sorts of customer characteristics, ranging from their demographic information (e.g., Age, Family) to whether they currently have various accounts open with the bank (e.g., Securities Account, CD Account). For a complete description of the fields, consult the data page on Canvas.

There are two characteristics of customers that would be used for the company to predict.

Task 1. First, the bank would like to predict a customer's Mortgage amount (the "Mortgage" variable.) The bank would like a model to make an educated prediction on income so that they may target customers with appropriate marketing.

Task 2. Secondly, the bank is aggressively trying to convert customers from depositors into borrowers through its personal loan program. The column **Personal Loan** shows whether each customer responded to a direct marketing campaign related to this program. The marketing team is now trying to understand what types of customers responds to new personal loans marketing. If they can establish reasonably strong predictive power, they will deploy the model more widely across their customer base to identify promising leads and a more nuanced target market.


# Task 1: Predicting Mortgage Amount

## Question 1

Do the following:

* Import the data as "bank"

* Remove all non-numeric variables from the data frame. Anything that feels more categorical than numeric should be nulled out. This is because in kNN, we cannot use any variable as predictors when we have difficulty calculating the distance. I would also say that anything that doesn't make sense as a numeric value (like Zip Code), or something that should not be predictive (like ID) of the dependent variable, should also be nulled out.


```{r}
library(caret)

Bank <- read.csv('UniversalBank.csv')

Bank$ID <- NULL
Bank$ZIP.Code <- NULL
Bank$Education <- NULL
Bank$Family <- NULL
Bank$Personal.Loan <- NULL
Bank$Securities.Account <- NULL
Bank$CD.Account <- NULL
Bank$Online <- NULL
Bank$CreditCard <- NULL


# Set a seed of 72 and partition a training set with 55% of the data.


# Set the seed for randomization
set.seed(72)

# Store the number of observations
N <- nrow(Bank)
# Round the number of observations we need for training set, which will be 55% of the observations.
trainingSize  <- round(N*0.55)
# Randomly sample (using the sample() function) to select the number of cases we need (trainingSize) from N cases total.  
trainingCases <- sample(N, trainingSize)
# Note that trainingCases now stores the rows that were randomly sampled.
#View(trainingCases) # You can see them here.

# Now let's divide the df into training and test cases
# Here, we subset the df based on the training cases
training <- Bank[trainingCases,]
#View(training) # Let's look at results!

#Then, we subset the df based on everything NOT in the training cases.
test <- Bank[-trainingCases,]
#View(test) # Let's look at the results!
```


## Question 3

* Build a kNN regression model using standardized features ("indepedent variables") to predict Mortgage in R. Set it up so that you are using the four closest neighbors for the predictions.

```{r}
# For standardized features
model <- knnreg(Mortgage ~ ., data=training, k=4, preProcess=c("center","scale"))
```

## Question 4

Apply the model to the test data frame. Then, store the predicted values in both the object "predictions" and within the test data frame so you can clearly see what the predictions are when you view the data frame.

```{r}

predictions <- predict(model, test)

test$predictions <- predict(model, test)
```

## Question 5
Evaluate the model. Calculate the MAPE and the RMSE. What are they? Interpret each of the them.

```{r}
# Store the observed Mortgage values
observations <- test$Mortgage

# Get the errors between observed and predictions (so these are "residuals"")
errors <- observations - predictions

#Calculate the MAPE (Mean Absolute Percentage Error) for the model:
mape <- mean(abs(observations-predictions)/observations)
mape

#Calculate the Root Mean Square Error (RMSE)
rmse <- sqrt(mean((observations-predictions)^2))
rmse

```

## Question 6
In one sentence, interpret the MAPE from the last question using the appropriate units. (Note: It is possible that you received a MAPE that is NaN or Inf, or basically incalculable. If you do, I will provide extra credit IF you can *clearly* explain why it is incalculable. Try to think about the MAPE equation and also take a look at the data.))

Answer: The MAPE is NaN as some of the observations are 0 and when R executes the formula, the 0 results in a value that is undefined, as dividing any value leads to an undefined value.

## Question 7
In one sentence, interpret the RMSE from Question 6 using the appropriate units.

Answer: RMSE is 111.6946, which suggests that the actual values and the observed values have an average deviation of $111685 as the units of the target variable is in $ thousands.

## Question 8
Calculate the benchmark MAPE and RMSE when using the mean as the prediction. Is your model useful? Why or why not?

Answer: The MAPE is Inf as some of the observations are 0 and when R executes the formula, the 0 results in a value that is undefined, as dividing any value with 0 is not possible. When comparing MAPE, the usefulness of the model cannot be determined as the answers are indefinite. 

In case of RMSE, the benchmark error is lower than the models error, which is 101.6465 (the model has an RMSE of 111.6946), indicating that the model is not useful as the error of model is higher than the benchmark.
```{r}
# Store the observed values subtracted by average of Mortgage
errors_bench <- observations - mean(training$Mortgage)

# What is the mape compared to benchmark
mape_bench <- mean(abs(errors_bench)/observations)
mape_bench
# This is about 0.308, or an average deviation of 30.8% when we just naively use the mean of the dependent variable as the benchmark.

rmse_bench <- sqrt(mean(errors_bench^2))
rmse_bench

```


# Task 2: Predicting Personal Loan Offer Acceptance

## Question 9

We will now turn our attention to creating a kNN Classification model to predict whether a customer would accept a personal loan offer. Make sure your code is in the following order:

* First, let's clear the global environment to make sure we don't confuse our previous data and model with this one.

* Second, Import the data as "bank" once again.

* Third, remove all non-numeric variables from the data frame except the target variable. Anything that feels more categorical than numeric should be nulled out. This is because in kNN, we cannot use any variable as predictors when we have difficulty calculating the distance. I would also say that anything that doesn't make sense as a numeric value (like Zip Code), or something that should not be predictive (like ID) of the dependent variable, should also be nulled out.

* Fourth, convert the target variable to a factor

* Fifth, standardize all the numeric predictors.

* Sixth, set a seed of 72 and partition a training set with 55% of the data once again.

```{r}

rm(list = ls()) # clears global environment
knitr::opts_knit$set(root.dir = 'C:/Users/USER/OneDrive - Institute of Business Administration/Desktop/QTM 666')

Bank <- read.csv('UniversalBank.csv')

Bank$ID <- NULL
Bank$ZIP.Code <- NULL
Bank$Education <- NULL
Bank$Family <- NULL
Bank$Securities.Account <- NULL
Bank$CD.Account <- NULL
Bank$Online <- NULL
Bank$CreditCard <- NULL

Bank$Personal.Loan <- as.factor(Bank$Personal.Loan)

library(caret) # preProcess() function requires package 'caret'.
# standardize the inputs to put variables on comparable scales
standardizer <- preProcess(Bank, c("center","scale")) # standardize the variables
Bank <- predict(standardizer, Bank)

# Setting seed and creating a training set with 55% of data.

set.seed(72)
N <- nrow(Bank)
trainingSize <- round(N*0.55)
trainingCases <- sample(N, trainingSize)
train <- Bank[trainingCases, ]
test <- Bank[-trainingCases, ]

```

## Question 10

* Train the kNN classification model using all available numeric inputs, using the four closest neighbors to make a prediction. Make sure to standardize your predictors.

```{r}
library(caret)
model <- knn3(Personal.Loan ~ ., data=train, k=4)

#Make predictions based on the model using test.
predictions <- predict(model, test, type="class")

#We can also store the predictions so we can see it.
test$predictions <- predict(model, test, type="class")

# Store the observed values
observations <- test$Personal.Loan
```


## Question 11

Create the confusion matrix to see the errors. 

```{r}
# confusion matrix
table(predictions, observations)
```

## Question 12

How many total predictions did the model get correct? (Not percentage)

Answer: A total of 2018 predictions were made correctly by the model.

```{r}
(1946+72)
```


## Question 13

How many total predictions did the model get incorrect? (Not percentage)

Answer: A total of 232 predictions were made incorrectly by the model.

```{r}
(92+140)
```


## Question 14

Manually calculate the error rate (according to numbers you received in the confusion matrix. Show calculations using R as a calculator.

Answer: 

```{r}
(92+140)/(1946+140+92+72)
```



## Question 15

Now, Calculate the error rate using R code. Make sure your manual calculation is correct.

```{r}
# Store the observed values
observations <- test$Personal.Loan

error_rate <- sum(predictions != observations)/nrow(test)
#error_rate = 1 - sum(predictions == observations)/nrow(test)   # same as above, but just different way to calculate.
error_rate
```


## Question 16

Calculate the benchmark error rate. You can do this using code. (NOte that, however- for practice- you should also be able to get the same benchmark error rate using the confusion matrix!)

Answer:

```{r}
# benchmarkErrorRate lives inside BabsonAnalytics.R
source('BabsonAnalytics.R')  # this adds the bechmarkErrorRate() function to your list of functions since there is no easy package for this.
error_bench <- benchmarkErrorRate(train$Personal.Loan, test$Personal.Loan)
error_bench
# the benchmark error rate is when you use the mode to make predictions, rather than the model.


#Manual Calc. of benchmark error
(140+72)/(1946+140+92+72)
```


## Question 17

Is your error rate for the model better than the benchmark?

Answer: In this case, the benchmark error is lower than the models error, which is 0.0942 (the model has an error of 0.1031), indicating that the model is not useful as the error of model is higher than the benchmark.


## Question 18

Calculate the sensitivity. You can use either code or the confusion matrix to do so manually.

Answer: 

```{r}
# R Code
sensitivity <- sum(predictions == 1 & observations == 1)/sum(observations == 1)
sensitivity

# Manual
sensitivity_manual <- 72/(140+72)
sensitivity_manual

```


## Question 19

Calculate the specificity. You can use either code or the confusion matrix to do so manually.

Answer:

```{r}
# R Code
specificity <- sum(predictions == 0 & observations == 0)/sum(observations == 0)
specificity

# Manual
specificity_manual <- 1946/(1946+92)
specificity_manual

```

